Evaluating infectious disease forecasts with allocation scoring rules

### 2023-07-10 alloscore_manuscript/allocation.tex
Continue allocation score discussion in intro.

>>>>

\subsection{Allocation Scoring}

In a nutshell, the allocation score of a multivariate forecast $F$ is the unnecessary realized unmet demand when $F$ is used to allocate so as to minimize expected unmet 
need --- that is, when we distribute a limited amount $K$ of our resource so that total expected unmet need according to $F$,
\begin{align}
\sum_i E_{F_i}[\text{unmet need in location $i$}]
\end{align}
is minimized, where the $F_i$ are the marginal distributions of $F$. By ``unnecesary'' we mean the unmet need that could have been avoided by an oracle that knows exactly how much need will occur in each location and divides the amount $K$ so that nothing is wasted in one location while it could be put to use in another.
This stochastic allocation problem has an intuitively appealing solution: allocate so that the probabilities of need exceeding allocation 
in various locations are as close to each other as possible (see methods).  This will lead to the allocations provided by $F$ being quantiles of $F_i$ for some \emph{single} probability level $\tau$.

For example, suppose we have a forecast $F$ for need in two locations with $F_1 = \mathrm{Unif}(0,8)$ and $F_2 = \mathrm{Unif}(4,8)$. If we have $K = 10$ units
of our resource available, the optimal allocation according to $F$ would be 4 units in location 1 and 6 units in location 2. This allocates to both locations at their forecast medians.  If, on the other hand, we only have $K = 3$ units available, we will necessarily have to allocate in location 2 at a quantile for $\tau = 0$, and therefore all 3 would go to location 2 since allocating any resources to location 1 would lead to unbalanced probabilities of need exceeding allocation. 
\footnote{Matching probabilities will not always be possible if we allow point masses in the $F_i$.  For example, if $F_2$ were the point forecast of 8 interpreted as a degenerate probabilistic forecast with mass 1 at 8, the optimal allocation for $K = 10$ would be 2 in location 1 and 8 in location 2. Notice though that both allocations are still technically medians.}

Next suppose that we observe needs of 8 and 3 in locations 1 and 2, respectively. With $K = 10$ units of the resource, the oracle would be able to prevent all
but 1 of the total 11 units of needs from going unmet. The forecast $F$'s allocation results, at the same time, in $8 - 6 = 2$ units of unmet need in location 1 despite leaving no need unmet in location 2. The allocation score for the forecast when $K = 10$ would therfore be
1 (= 2 realized $-$  1 unavoidable) in units of unnecessary unmet need.  When $K = 3$, however, the forecast does not ``over-allocate'' in either location which leads to the same unmet need of $11 - 3 = 8$ as does the oracle's allocations.  The allocation score for $F$ when $K = 3$ is therefore 0.

Note the way in which $F$ incurred a positive (i.e., non-optimal) allocation score of 1 for $F$ when $K = 10$. It was not directly directly due to individual misalignments 
of the marginal forecasts $F_i$ with the observed needs, but rather because the allocations and observed needs were not proportional as vectors.
This illustrates a fundamental property of the allocation score.  For a given probability level $\tau$, if the quantiles $Q_i(\tau)$ of the marginal forecasts
are proportional to the observed needs $y_i$, then the allocation score is zero for the resource constraint level $K = \sum Q_i(\tau)$. This stands in marked constrast to other common scoring methods for multi-variate forecasts that aggegate univariate scores such as CRPS or WIS for the marginal forecasts where a 
misalignment in one coordinate is penalized regardless of alignments in other coordinates. \apg{Should this be illustrated here, or elsewhere?}

It is therefore often straighforward to constuct forecasts $F$ and $\widetilde{F}$ for a given outcome distribution that switch rankings under the allocation and traditional scores by ensuring that the marginal forecasts of $F$ center sharply around allocations that are proportionally similar to the central tendencies of the outcome distribution but are strongly biased.

While multivariate scoring rules have not seen wide application in infectious disease forecast evalution, it seems that the allocation score would have a similar relationship with them, since while bias penalties in multi-variate scores can be offset by better forecasting of dependence structure, the proportional biases that allocation scoring are insensitive to are not \emph{per se} tolerated by multivariate scoring rules such as the energy score, variogram score, or Dawid-Sebastiani score.
